---
title: "Analysis of COVID-19 Pandemic Related Tweets"
author: | 
  | 10867436
  | 10886338
  | 10835879
  | 10760993

date: "5th December 2023"
output: beamer_presentation
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Table of Content
- Introduction
- Description of dataset
- Information about the users
- Analysis of Tweets Pattern
- Analysis of Word Usage in Tweets
- Analysis of Users Sentiment in Tweets
- Analysis with T-test
- Limitations of the dataset
- Conclusion & Recommendations
- References

# Introduction
As the world grappled with the challenges posed by the pandemic, social media platforms became a significant avenues for individuals to share their thoughts, concerns, and experiences.


These tweets were analysed to provides insights into the sentiments expressed during the unprecedented times of the COVID-19 pandemic.


# Description of Dataset
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Set working directory
#setwd("C:/Users/adewu/OneDrive - University of Plymouth/Desktop/Math 513/Math_513_Presentation_Dec_7_2023")

# Load the required libraries
suppressMessages({
library(tidyverse);
library(tidytext);
library(textdata);
library(igraph);
library(tm);
library(ggraph);
library(devtools);
library(scales);
library(ggpubr);
library(RColorBrewer)
library(ggwordcloud)
  })

# Load the dataset in R
corona_data <- read_csv("Corona_NLP_train.csv", 
    col_types = cols(TweetAt = col_date(format = "%d/%m/%Y")))

# get dimension of the dataset
dim_data <- dim(corona_data)
```

The dataset is made up of tweets that were extracted from Twitter now called `X`.

The dataset contains tweets about the Covid-19 pandemic between the period of March and April 2020 from `r label_comma()(dim_data[1])` users.

It contains information used to identify the users such as their locations, and the time of tweet and their original tweets.

# Information about users
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}
# Provide summary statistic about unique users
unique_users <- corona_data %>% 
                        distinct(UserName) %>% 
                        nrow()
unique_location <- corona_data %>% 
        distinct(Location) %>% nrow()
```

A total of `r label_comma()(unique_users)` users from `r label_comma()(unique_location)` locations contributed to tweets in the dataset.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Checking the proportion of the dataset without location data
missing_percent <- sum(is.na(corona_data$Location))/nrow(corona_data)

# improved the the dataset by recoding countries code to country name
corona_data_rec <- corona_data %>% mutate(Location_rec = 
                      recode(Location, "London" = "United Kingdom",  
                        "USA" = "United States of America", 
                        "New York, NY" = "United States of America",
                        "New York" = "United States of America",
                       "New York City" = "United States of America",
                       "Atlanta, GA" = "United States of America",
                       "Chicago, IL" = "United States of America",
                       "Houston, TX" = "United States of America",
                       "Los Angeles, CA" = "United States of America",
                       "UK" = "United Kingdom",
                       "Washington, DC" = "United States of America",
                       "Washington, D.C." = "United States of America",
                       "San Francisco, CA" = "United States of America",
                       "Boston, MA" = "United States of America",
                       "Chicago, IL" = "United States of America",
                       "Toronto, Ontario" = "Canada",
                       "Austin, TX" = "United States of America",
                       "Toronto" = "Canada",
                       "Mumbai" = "India",
                       "Dallas, TX" = "United States of America",
                       "London, England" = "United Kingdom",
                       "United States" = "United States of America",
                       "California, USA" = "United States of America",
                       "New York, USA" = "United States of America",
                       "New Delhi, India" = "India",
                       "Las Vegas, NV" = "United States of America"
                    ))
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Visualization of the top five locations by tweets count
corona_data_rec %>%
  count(Location_rec, sort = TRUE) %>% # count the frequency for each location
        na.omit()%>% 
  mutate(Location_rec = reorder(Location_rec, n)) %>% # make sure that locations are ordered according to frequency
  head(5) %>%
  ggplot(aes(x = Location_rec, y = n, fill = Location_rec)) +
  geom_col(show.legend = FALSE) +
  coord_flip() + # flip x and y axes
  labs(x = "Top 5 Locations",
       y = "Frequency",
       title = "Top 5 User Locations in Covid-19 Related Tweets") + 
  theme(axis.text = element_text(size = 14, color = "black"), 
        axis.title = element_text(size = 14, color = "black"),
        title = element_text(size = 16))
```


# Analysis of Tweets Pattern
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Time series visualization to understand the rate of tweets across different times
# Days with highest number of tweets about corona virus
corona_data_rec %>%
  count(TweetAt) %>%
  ggplot(aes(x=TweetAt, y=n)) +
  geom_bar(stat="identity", fill="thistle2", colour="black") +
  theme_bw() +
  labs(x="Date (day)", y="Frequency",
       title = "Changes in Tweets Frequency by Day")+
        theme(axis.title = element_text(size = 15),
              axis.text = element_text(size = 14),
              plot.title = element_text(size = 16))+
        scale_y_continuous(labels = label_comma())
```


# Analysis of Word Usage in Tweets
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Creating a function for text preprocessing
cleantweets <- function(text) {
  text %>%
    # Remove URLs
    str_remove_all(" ?(f|ht)(tp)(s?)(://)(.*)[.|/](.*)") %>%
    # Remove mentions e.g. "@my_account"
    str_remove_all("@[[:alnum:]_]{4,}") %>%
    # Remove hashtags
    str_remove_all("#[[:alnum:]_]+") %>%
    # Replace "&" character reference with "and"
    str_replace_all("&amp;","") %>%
    # Remove punctuation, using a standard character class
    str_remove_all("[[:punct:]]") %>%
    # Replace any newline characters with a space
    str_replace_all("\\\n", " ") %>%
    # Remove the digit "2"
    str_remove_all("[:digit:]") %>%
        # Remove any trailing whitespace around the text
    str_trim("both") %>%
    # Make everything lowercase
    str_to_lower()
    
}

```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# we are creating a new column `tweet_loc` from our location and storing as character for later use in the analysis.
corona_data_rec$tweet_loc <- as.character(corona_data_rec$Location_rec)

# Tokenization of text data, `OriginalTweet` in `Corona_data_rec` using clean tweet function
# cleaning up the Original tweet column with cleantweets function
corona_data_rec$cleantweet <- cleantweets(corona_data_rec$OriginalTweet)

clean_corona <- corona_data_rec %>%
  select(UserName, tweet_loc, cleantweet, TweetAt) %>%
  unnest_tokens(word, cleantweet)

# Define our own stopwords that we don't want to included in our data
our_stop_word <- data.frame(word = c("covid"))

# remove our own stopwords from the list of words too
data("stop_words")
clean_corona_2 <- clean_corona %>%
        anti_join(stop_words) %>%
        anti_join(our_stop_word)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width="90%"}
# Visualizing most frequently used words in the tweets
plot1 <- clean_corona_2 %>%
  count(word, sort = TRUE) %>% # count of number of occurencies of each word and sort according to count
  head(10) %>% # extract top 10 words
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n, fill = "mediumblue")) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(x = "Unique Words",
       y = "Frequency") +
  theme(axis.text = element_text(size = 14, color = "black"), 
        axis.title = element_text(size = 14, color = "black"),
        title = element_text(size = 12))+
        scale_y_continuous(labels = label_number(scale = 1e-3,suffix = "k"))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Aggregating data to visualize top words by frequency
clean_corona_3 <- clean_corona_2 %>%
  count(word, sort = TRUE) %>% 
  mutate(freq = n / sum(n))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# Perform With the package wordcloud
plot2 <- ggwordcloud(clean_corona_3$word, clean_corona_3$freq, 
               min.freq = 1, 
               max.words = 100,
               random.order = FALSE, 
               colors = brewer.pal(8, "Dark2"), 
               scale = c(3, 0.5),
               rot.per = 0.1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#creating a grouped plot to show wordcloud and bar chart
figure_4_topwords <- ggarrange(plot1, plot2, ncol = 2, align = "hv",
                               common.legend = TRUE)
#adding annotatons to the visual
annotate_figure(figure_4_topwords,
        top = text_grob("Most frequently used words in Covid-19 related tweets",
                        color = "grey51", face = "bold", size = 14))
```


# Analysis of Users Sentiment in Tweets
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}
# The bing lexicon categorizes words into positive and negative categories in a binary approach.
# Join sentiment classification to the tweet words
bing_word_counts <- clean_corona_2 %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE, echo=FALSE}
# Visualizing the 10 most common tweets words for both Negative and Positive words in the corona related tweets displaying the tone of the users sentiments.

bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.5, size = 3) +  # Added this line for displaying numbers
  coord_flip() +
  coord_flip() +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(title = "Most common Negative and Positive words in Covid-19 Related Tweets",
       y = "Sentiment",
       x = NULL) +
  theme(axis.text = element_text(size = 14, color = "black"), 
        axis.title = element_text(size = 14, color = "black"),
        title = element_text(size = 14),
        strip.text = element_text(size = 14))+
        scale_y_continuous(labels = label_comma())
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE, results='hide'}
# Calculating the sentiments score and separating them into `Positive` and `Negative` scores
corona_sentiment <- clean_corona_2 %>%
  inner_join(get_sentiments("bing")) %>%
  count(UserName, sentiment) %>%
  spread(sentiment, n, fill = 0) %>% # negative and positive sentiment in separate columns
  mutate(score = positive - negative) # score = net sentiment (positive - negative)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE, results='hide'}
# working out the mean sentiment score 
sentiment_means <- corona_sentiment %>% 
  summarize(mean_score = mean(score))
```


# Analysis of Users Sentiment in Tweets (cont.)
```{r message=FALSE, warning=FALSE, paged.print=FALSE, echo=FALSE}
ggplot(corona_sentiment, 
       aes(x = score)) + # Sentiment score on x-axis
  geom_bar(fill = "lightgreen", colour = "darkgreen") + # geom_bar will do the tabulation for you :-)
  geom_vline(aes(xintercept = mean_score), data = sentiment_means) +
  # Add a vertical line at the mean score, calculated and stored in sentiment_means_nba above
  geom_text(aes(x = mean_score, 
                y = Inf, 
                label = signif(mean_score, 3)), # Show to three significant figures
            vjust = 2,
            data = sentiment_means) + 
        scale_y_continuous(labels = label_comma())+
  # Add the mean as a number; just moves it down from the top of the plot
  scale_x_continuous(limits = c(-10,10), breaks = -10:10,  # Specify a suitable integer range for the x-axis
                     minor_breaks = NULL) + # Show integers; set this to a suitably large range
  labs(x = "Sentiment Score", y = "Number of tweets",
       title = paste("Sentiments towards the Corona virus give a mean of", signif(sentiment_means$mean_score, 3)))
       # Title that gives page name and mean sentiment score, to three significant figures
       
```

# Analysis of Users Sentiment in Tweets (cont.)
```{r message=FALSE, warning=FALSE, paged.print=FALSE, echo=FALSE}
# we are interested in creating a visualization to show the 10 most common used words from 4 top Location in the corona dataset.
# get the top 4 Locations from our data frame is named clean_corona_2
top_locations <- clean_corona_2 %>%
  filter(!is.na(tweet_loc)) %>%
  count(tweet_loc, sort = TRUE) %>%
  top_n(4, n) %>%
  pull(tweet_loc)

# Create a plot for the top 5 words in each of the top 4 locations
plot_data <- clean_corona_2 %>%
  filter(tweet_loc %in% top_locations) %>%
  count(tweet_loc, word, sort = TRUE) %>%
  group_by(tweet_loc) %>%
  top_n(5, n)

ggplot(plot_data, aes(x = reorder(word, n), y = n, fill = word)) +
  geom_col() +
  facet_wrap(~tweet_loc, scales = "free_y", ncol = 2) +
  labs(x = "Words",
       y = "Frequency",
       title = "Top 5 Words in Top 4 Locations") +
  theme(axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 12, color = "black"),
        title = element_text(size = 14)) +
  theme(legend.position = "none") +
  coord_flip()

```


# Bigram Analysis of Tweets Text
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE, results='hide', echo=FALSE}
# we are interested in text analyses that are based on the relationships between two words and two most common used words in this sentiment analysis.
# this function is to clean the data of all stop words.
count_bigrams <- function(dataset) {
  dataset %>%
    unnest_tokens(paired_words, OriginalTweet, token = "ngrams", n = 2) %>%
    separate(paired_words, c("word1", "word2"), sep = " ") %>%
    filter(!word1 %in% stop_words$word) %>%
    filter(!word2 %in% stop_words$word) %>%
    count(word1, word2, sort = TRUE)
}

corona_paired <- count_bigrams(corona_data)

# we still need to get rid of some unwanted words, with a custom stop word remover.
# Define our own stopwords that we don't want to include in our bigrams
ourstop_words <- c("https", "t.co")

# Filter out the chosen stop word
corona_paired_clean <- corona_paired %>%
  filter(!(word1 %in% ourstop_words) & !(word2 %in% ourstop_words))

# From data frame is named corona_paired_clean
top_20_bigrams <- corona_paired_clean %>%
  top_n(20, n)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE, echo=FALSE}
# Create a bubble plot for the top 20 bigrams
ggraph(top_20_bigrams, layout = "fr") +
     geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                    arrow = grid::arrow(type = "closed", length = unit(2, "mm")), 
                    end_cap = circle(1, "mm")) +
     geom_node_point(color = "lightblue", size = 2) +
     geom_node_text(aes(label = name), size = 4) +
     theme_void()

```

# Analysis with T-test

The Independent T-test is a parametric test that is used to find the mean difference in a quantitative variable across a categorical(dichotomous) variable i.e a variable with only two levels or categories. It also follows some other assumptions that data from the two groups must be more than thirty each and the responses must have been randomly selected.

Considering the fact that our data set contains quite a large sample size, there were no quantitative variables in the data set, likewise any categorical(dichotomous) variable with two possible outcome with which we could compare mean differences. Also the data weren't randomly selected and as such a T-test analysis was not applicable in this analysis.

# Limitations of the dataset
- The dataset was drawn from twitter that does not properly represent the general population.

- The temporal breadth of the dataset might not accurately reflect the emotion across a longer time frame, and it might be especially impacted by certain events that occur only in March and April 2020.

# Conclusion & Recommendations
During the early phases of the pandemic, people expressed shock, worry, and uncertainty  as evidenced in the sentiment analysis score of -0.6. They also expressed concerns about the virus's capacity to spread, the lack of sufficient information, and its possible effects on daily life and health. A dip in number of tweets were noticed between March 28th & 30th 2022; this may have been due to information overload, or distraction by other events which temporarily diverted focus.

In light of these findings, The following recommendations were proposed

- The government should set up clear and open channels of communication to stop the spread of false information and calm public anxiety. 
- Results showed the often-used negative terms by people were "panic" and "crisis." They should have confidence in the competence of authorities to handle emergencies, and ensure resources are well distributed.


# References

Csárdi, G., & Nepusz, T. (2006). The igraph software package for complex network research. InterJournal, Complex Systems, 1695. <https://igraph.org>

Feinerer, I., Hornik, K., & Meyer, D. (2008). Text Mining Infrastructure in R. Journal of Statistical Software, 25(5), 1-54. <https://www.jstatsoft.org/article/view/v025i05>

Le Pennec E, Slowikowski K (2023). _ggwordcloud: A Word Cloud Geom for 'ggplot2'_. Rpackage version 0.6.1, <https://CRAN.R-project.org/package=ggwordcloud>

Silge, J., & Robinson, D. (2016). tidytext: Text Mining and Analysis Using Tidy Data Principles in R. Journal of Open Source Software, 1(3), 37. <https://joss.theoj.org/papers/10.21105/joss.00037>

Wickham, H. (2019). tidyverse: Easily Install and Load the 'Tidyverse'. R package version 1.3.1. <https://CRAN.R-project.org/package=tidyverse>
